A Case Study in Preserving a High Energy Physics Application

### 2014-01-29 paper.tex

\section{Overview of Application}

Matthias, one graduate student from Physics Department of University of Notre
Dame, finished one interesting tau analysis. Others want to understand how the tau
analysis works and repeat it.

\subsection{Background}

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is too short-lived to be detected
itself, it has to be reconstructed from its decay products.  The
analysis discussed in this section focuses on the Higgs decaying to
two tau leptons, which are themselves reconstructed from particle
showers.  Within the analysis, events are identified that show a
signature of decay products compatible with both hadronic tau and top
decays.  Properties of such events are used to distinguish signal-like
(matching simulated Higgs decays) and background-like events and are
also used in further statistical analysis.

The CMS collaboration provides analysis end-users with a pre-processed
and reduced data format, AOD, containing information for events, i.e.,
proton-proton collisions with a signature of interest, in the form of
reconstructed particles.  This format is based on the RAW output of
the CMS detector readout electronics and reconstructed world-wide.
Both real and simulated data are available for examination.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=.8\columnwidth]{data-code-size.eps}
\caption{Inputs to Tau Analysis}
\label{fig:data-code-size}
\end{figure}

\begin{table}[t]
    \centering
    \small
    \begin{tabular}{|l|l|r|r|r|}
        \hline
        \bf Name & \bf Location & \bf Total & \bf Named & \bf Used \\ 
        \hline
        SBEAN data    & hdfs & 33TB & 43.3GB & 20GB \\ \hline
        CMSSW binaries     & local & 1.16TB & 449MB & 5.2MB\\ \hline
        Tau source       & git & 51MB & 51MB & 51MB \\ \hline
        PyYAML binaries    & http & 52MB & 51MB & 51MB \\ \hline
        .h file       & http & 41KB & 41KB & 41KB \\ \hline
        Misc commands & hpc fs & 206TB & ??? & 119MB \\ \hline
        Linux commands & local fs & ??? & ??? & 88.4MB \\ \hline     
        Total          &          & ??? & ??? & ??? \\ \hline
    \end{tabular}
    \caption{Data and Code Used by Tau Analysis}
    \label{table:size-original-real}
\end{table}

\section{Overview of Tau Analysis}

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is to short-lived to be detected
itself, it has to be reconstructed from its decay products.

The application which is the study of this paper is called \emph{TauAnalysis}
It searches for cases where the the Higgs boson decays to two tau leptons.
The leptons are not observed directly, but by the particle showers
that they generate.  So, the analysis must search for detector
events that show a signature of decay products compatible with both hadronic tau and top decays.  Properties of such events are used to distinguish
the events of interest (Higgs decays) from all other events and
are also used in further statistical analysis.

Figure~\ref{fig:data-code-size} shows that both the code and data
that form \emph{TauAnalysis} are drawn from large repositories through
multiple steps of reduction.

{\bf Data Sources.}
The CMS collaboration provides analysis end-users with a pre-processed
and reduced data format, AOD, containing information for events, i.e.,
proton-proton collisions with a signature of interest, in the form of
reconstructed particles.  This format is based on the RAW output of
the CMS detector readout electronics and reconstructed world-wide.
Both real and simulated data are available for examination.
